When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-01-25 doc/paper/vpd_et_paper_PG_v1.tex
just have discussions section, conclusions, and thenr ead forward

$\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and $\Delta$. The psychrometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the slope of the Clauisu-Clapeyron relationship $\Delta$. This is as expected, given that the aerodynamic conductance represents the efficiency fo the transfer of the surface environment to the atmosphere. As aerodynamic conductace  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).

$\Delta$'s presence in the scaling term also matches physical intuition. Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with decreasing temperature. The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$). So as $\Delta$ increases, you will get a larger damping of ET due to evaporative cooling.  The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly. In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness. So most of the differences in scaling between PFT should be in the aerodynamic conductance term. One interesting side note is that the coefficient of variability for both aerodynamic conductance and the scaling term is relatively constant across PFT, suggesting that the influence of aerodynamic conductance on the relative (to the PFT mean) variability of the scaling term is approximately similar across PFT.

Figure \ref{scale_vary}A shows the scaling term normalized by mean aerodynamic conductance (calculated for each plant functional type), and confirms that much of the relative variability of the scaling term is caused by aerodynamic conductance variability. Generally, temperature (via $\Delta$) causes less relative variability. However, the impact of $T$ on the relative variability increases with increasing aerodynamic conductance.

While the relative variability of scaling term is similar across PFT, the absolute value of scaling term varies strongly across PFT. Figure \ref{scale_vary}B shows scaling term evaluated at the mean aerodynamic conductance for each PFT, and at the range of observed temperatures for each PFT. As expected, for the tree PFTs (DBF, ENF) the scaling term is much larger and the temperature dependence is much stronger. Systematic differences in observed temperatures also cause differences in the average magnitude of scaling term. For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term. Additionally, because the variability of aerodynamic conductance increases proportionally to the mean, the spread of the scaling term due to aerodynamic conductance variability will be larger for the tree PFTs, although this is not shown for simplicity. To summarize Figure \ref{scale_vary}: the variability of the scaling term within each PFT will look like Figure \ref{scale_vary}A for each PFT, but the scale of the y-axis will increase or decrease according to mean aerodynamic conductance observed in Figure \ref{scale_vary}B.

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{./fig04.pdf}
\caption{Primary sources of variability for scaling term. A) Variability within each PFT: scaling term normalized by mean $g_a$ for each PFT. B) Variability between each PFT: scaling term evaluated at mean $g_a$ for each PFT. Temperature range is 5-95th percentile for each PFT. Additionally, stars denote the location of the 25th, 50th, and 75th percentiles.}
\label{scale_vary}
\end{figure}

%%%%% replace below with idea:
\subsection{Bulk statistics of $\frac{\partial \; ET}{\partial \; VPD}$}

Table 3 confirms our expectations for PFT behavior of $\frac{\partial \; ET}{\partial \; VPD}$. For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; VPD}$ is less than zero. However, $\frac{\partial \; ET}{\partial \; VPD}$ evaluated at the average of all variables (e.g. $\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA. And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; VPD}$ < 0 less than half the time. These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; VPD}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.

The units of $\frac{\partial \; ET}{\partial \; VPD}$ make it difficult to interpret if $VPD$ is even a meaningful contributor to ET's variability. To better understand $VPD$'s contribution, we multiply $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ with $VPD$'s standard deviation to define a (linearized) relative change in ET for variations in $VPD$ . $VPD$'s contribution to ET's variability ranges between 30 - 40 W m$^{-2}$ for all PFTs except for CSH, which is about 100 W m$^{-2}$. Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} \cdot std(R)$, as net radiation is generally the driver of ET (cite joe berry here)\explain{Find citation?}. For all PFTs except for CSH $VPD$ contributes between 30 - 40 \% of $R$'s contribution to variability. For CSH the portion is much larger, about 88 \%. $VPD$'s variability is certainly a non-negligible contributor to $ET$'s variability.

Theoretical derivation has so far illuminated the dependence of $\frac{\partial \; ET}{\partial \; VPD}$ and how it varies across PFT. Large mean $uWUE$ shifts CRO and DBF towards a tendancy to increase ET with VPD (positive $\frac{\partial \; ET}{\partial \; VPD}$), but a lower slope parameter ($g_1$) for DBF allows for decreaes in ET with VPD, at low environmental VPD. \explain{physically state how slope parameter is trade for water at leaf level, but uWUE is trade-ff between GPP and VPD at ecosystem-level}.  ENF's very low slope parameter ($g_1$), which is even lower than for DBF, increases the dependence of ET response  ($\frac{\partial \; ET}{\partial \; VPD}$) on $VPD$, and makes the function strongly nonlinear. This has the effect of making mean ET response ($\overline{\frac{\partial \; ET}{\partial \; VPD}}$)  more negative for a given frequency of occurence of negative ET response ($\frac{\partial \; ET}{\partial \; VPD} < 0$), than for other PFT. GRA shows the opposite behavior; a relatively high slope parameter ($g_1$)  makes the function more linear, causing a more positive mean ET reponse ($\overline{\frac{\partial \; ET}{\partial \; VPD}}$) for a given given frequence of negative ET reponse  ($\frac{\partial \; ET}{\partial \; VPD} < 0$). Finally, CSH's very low $uWUE$ dominates the slope parameter's effects, causing ET to almost always decrees with increasing atmospheric dryness (VPD), even with uncertainty included. Variability in $VPD$ accounts for the largest amount of $ET$ variability for CSH. For the other PFTs, $VPD$ contributes less to $ET$ variability, but still represents about 30-40 \% of $R$'s contributions to ET variability. \explain{think pierre is correct, could probably scrap this anecdote: WHAT IS THE SORUCE OF VARIBILITY OF THE REST OF ET I DON"T KNOW IF THAT IS RELEVANT FOR THE PRESENT PAPER.}

\begin{table}
\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT. \textbf{***dated, need to update these statistics***}}
\centering
\begin{tabular}{l c c c c c}
  \hline
PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
  \hline
CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
\hline

@@

\subsection{Validation of theory at eddy-covariance sites}
\label{testing}
So far we have developed a theory for $\frac{\partial \; ET}{\partial \; VPD}$'s behavior as a function of VPD and PFTs. We now turn to an direct evaluation against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}).

Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict. Our goal was to capture the leading order behavior of the ET dependence on VPD. Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a large amount of noise when reprdrocuing the derivatives of ET.

For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBG figure looks as one would expect if one just added noise on top of our theory. The VPD dependence, given by the slope parameter ($g_1$), floows the median values of each box plot. Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations. Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves. The uncertainty is non negligable; there are bany observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.

@@

The theory does not match as well for CRO as for DBF and CSH. At low VPD the VPD dependence (due to $g_1$) is similiar to the leading order behavior of the observations, but the entire curve is shifted towards postiive values, due to some error in $uWUE$. Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependnce, is behavior that is impossible fo the theory to capture. This divergence between the theory and the observations for CRO could have been expected. For practical reasons, our theory does not include important sources of variability (particularly between sites) for CRO. We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites and/or years. This deficiencies would explain the inability of our theory to match obsercations. For example, a superpositions of sites with C3 plants (low VPD) and C4 plants (high VPD) would explain observed shift back towards negative ET reponse at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.

The theory for GRA suffers from similiar limitations as for CRO. The differences between the theory and the observations are more pronounced for GRA, but are explainable using similiar logic as that used for CRO. GRA will also have differences between sites in C3 vs C4 photosythesis, and irrigation practices, as well as strong seasonal variability in crop height and LAI. For GRA, the observations show a very different VPD dependence of response than the theory, with a strong tendency back toewards a negative ET reponse at high VPD which our theory cannot account for. As with CRO, this could be explained by a superposition of sites with C3 vs C4 sites. Unfortunately, because detailed site-level data of C3 vs C4 plant type, as well as irrigation practes, was difficult to obtain, we could not bin GRO and GRA further by irrigation and photosyethesis stragtegy. However, we hypothesize that the theory would much more closel match obsercations with photosythesis and irrigation are accounted for.

The comparison between the theory and obsercations for ENF is unique from the rest of the PFTs. The VPD dependence of the theory, determined by the slope paramter, is not well captured by observations. However, the theory closely captures the x-intercept and VPD$_{crit}$. So, the theory still captures when the ET reponse will be positive or negative, but is less accurate with the magnitude of the reponse. Therefore the theory is still useful for addressing when VPD drives or reduces ET. Also, it's with noting that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence. We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori physical constants whenever possible.

While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how would using PET, as many hydrometeorological studies and applications do, compare with our theory.  Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5. In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far more negative than 3.5 for the observations, even with all uncertainty included. So the PET reponse to atmospheric dryness is outside the range of all observable responses. This casts doubt on PET's usefulness for capturing the leading order behavior of water budgeting at the land surface. However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ as an alternative to PET to estimating land response and moisture budgeting.

\subsubsection{Testing VPD$_{crit}$}
As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of  $\frac{\partial \; ET}{\partial \; VPD}$ is determined. When $VPD < VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD}  < 0$ (plant response dominates), and when $VPD > VPD_{crit}$, $\frac{\partial \; ET}{\partial \; VPD} > 0$ (atmospheric demand dominates). Theoretically this $VPD_{crit}$ is only a function of PFT. We can look at how this value of $VPD_{crit}$ matches the data as different observed variables (including uncertainty) are allowed to vary.

Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty. Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT. Vertical black lines denote where our theory says $VPD_{crit}$ should be. For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates). In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.

The scatter figures generally confirm expectations from Figure \ref{agu_fig}. By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$. $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs. $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}. However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible. The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).

While $g_a$ has a large impact on the magnitude of the scaling, temperature, through the derisive of the Clausius-Clapeyron, $\Delta$, term, does not appear to have as much of an impact in the complete estimation of $\frac{\partial \; ET}{\partial \; VPD}$. Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance. Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer. While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling $\frac{\partial \; ET}{\partial \; VPD}$.

In general Figures \ref{real} and \ref{agu_fig} match our expectations, even with the large sensitivity to $\sigma$ seen in Figure \ref{term3}. Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO. We therefore have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.

>>>>

$\frac{P}{T}$ is an air-density term, which we would expect to vary relatively little compared to aerodynamic conductance and $\Delta$. The psychrometric constant ($\gamma$) is also relatively constant, so the scaling term should be primarily a function of aerodynamic conductance and temperature, through the Clauisu-Clapeyron relationship $\Delta$. This is as expected, given that the aerodynamic conductance represents the efficiency of exchange between the surface and the atmosphere. As aerodynamic conductace  increases, any plant response will be communicated more strongly to the atmosphere (and vice-versa).

$\Delta$'s presence in the scaling term also matches physical intuition. Evaporative cooling will dampen the ability of the atmosphere to take more moisture, because $e_{s}$ decreases with decreasing temperature. The decrease in $e_{s}$ is proportional to $\Delta$ ($\delta e_{s} = \Delta \delta T$). So as $\Delta$ increases, you will get a larger damping of ET due to evaporative cooling.  The functional from of $\Delta$ will be the same across PFT, but the temperature range may vary slightly. In contrast, aerodynamic conductance will vary strongly with PFT due to the importance of surface roughness for aerodynamic conducatance. So most of the differences in scaling between PFT should be in the aerodynamic conductance term.

The control of scaling term variability between PFTS by aerodynamic conductance is confirmed by the data (Figure \ref{scale_vary}). Differences between PFT are almost entiredly due to differences in aerodynamic conductance, rather than differences in observed temperature ranges. The Scaling term for the tree PFTs (DBF, ENF) is generally more than double the scaling terms for the other PFTs which have lower surface roughness and gererally smaller aerodynamic conductance.

Within each PFT, scaling term variability is controlled both by environmental temperature and aerodynamic conductance variability (Figure \ref{scale_vary}). While the observed variability of aerodynamic conductance conrbutes more to scaling term variability than temperature, temperature's contributions are non-negliable. Specifically, the scaling term is generally larger at low temperatures when ET will be weakly damped by Clasius-Clapyeron, and the effect of aerodynamic conductance variability is stronger as it is less weakly damped by Clasius-Clapyeron.

To summarize, variability between PFTs is mostly controlled by systematic differences in aerodynamic conducntace, due to differences in surface roughness, betwen each PFT. In contrast, variability within PFT is also controlled by temperature, through Clasius-Clapyeron. But, aerodynamic conductance variability generally impacts the scaling term more than temperature, even within PFT.

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{./fig04b.pdf}
\caption{Primary sources of variability for the scaling term, as a function of PFT. Note the change in axis scale for each plot. The 5th-95th percentile range of temperature is plotted at the 5th, 25th, 50th, 75th, and 95th percentiles of aerodynamic conductance, as observed for each PFT's sites.}
\label{scale_vary}
\end{figure}

\subsection{Bulk statistics of ET response to VPD}
\label{stats_sec}
Previous sections only considered idealized scenarios without uncertainty ($\sigma = 1.0$). In this section we will consider bulk statistics, as a function of PFT, of ET reponse when $\sigma$ is allowed to vary and eddy-covariance derived uncertainty is included (Section \ref{methods}).  These bulk statistics of ET reponse confirm our expectations based on the theory presented in Sections \ref{sign_term}-\ref{scale_term} (Table \ref{stats}. For all PFTs, mean ET reponse to increasing VPD is negative. However, ET response evaluated at the average of all variables (e.g. $\sigma$, $T$, $c_s$, $VPD$) is only negative for CSH and GRA. This difference in mean ET reponse as compared to the ET reponse at mean environmental conditions is due to the non-linear nature of the reponse, in which negative reponses are generally much larger magnitude than postive reponses (Figure \ref{sign}).

Regarding the frequency of negative and postive ET response, observations of ENF, GRA and CSH are more frequently characterized by a negative ET reponse to increasing VPD, and are thus more frequently water conservative. DBF and CRO observations have greater frequency of postive ET reponse to increasing VPD, so are less frequently water conservative.

In general the bulk statistics match our expectations based on previous idealzied and theoretical analysis, with the possible exception that inclusion of uncertainty shfits grass towards a slightly more negative reponse than the theory and Figure \ref{sign} would suggest. The bulk statistics motivate a more throrough examination of the structure of uncertainty and more sophsiticated validation of our theory's performance against observations.

\begin{table}
  \label{stats}
\caption{Statistics of $\frac{\partial \; ET}{\partial \; VPD}$ as a function of PFT. \textbf{***dated, need to update these statistics***}}
\centering
\begin{tabular}{l c c c c c}
  \hline
PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)$ & $\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)$ & $\frac{\frac{\partial \; ET}{\partial \; VPD}\left(\overline{env}\right)*\text{std}(VPD)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{env}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
  \hline
CRO & 0.000853 & 0.026241 & 37.05 & 0.41 & 0.473311\\
CSH & -0.108234 & -0.091526 & 101.72 & 0.88 & 0.931660\\
DBF & -0.012727 & 0.013794 & 39.47 & 0.33 & 0.461674\\
ENF & -0.034087 & 0.000706 & 33.22 & 0.30 & 0.534425\\
GRA & -0.019637 & -0.000921 & 33.60 & 0.35 & 0.631735\\
\hline

@@

\subsection{Validation of theory at eddy-covariance sites}
\label{testing}
We now turn to an direct evaluation of the functional forms presented in Section \ref{sign_term} against observations taken from a wide range of eddy-covariance datasets (Section \ref{methods}). Figure \ref{agu_fig} shows the distribution of of the sign term (when uncertainty is included), as compared to what the theory would predict. Our goal was to capture the leading order behavior of the ET dependence on VPD. Given the assumptions we have made, and the uncertainties of flux tower observations themselves, we expect a large amount of noise when reprdrocuing the derivatives of ET.

For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included, and the DBG figure looks as one would expect if one just added noise on top of our theory. The VPD dependence, given by the slope parameter ($g_1$), follows the median values of each box plot. Perhaps most importantly, the x-intercept, and thus VPD$_{crit}$, matches nearly exactly between the theory and the observations. Therefore the sign of the ET response to increases in VPD should be well matched, subject to the unavoidable constraints of noise, much of which may be from the observations themselves. The uncertainty is non negligable; there are many observations in each box for which the the sign of observation is opposite the response predicted by the theory, but to leading order our theory matches the observations.

@@

As Section \ref{stats_sec} suggested, the theory does not match as well for CRO as for DBF and CSH. At low VPD the VPD dependence (due to $g_1$) is similiar to the leading order behavior of the observations, but the theory's curve is shifted towards postiive values relative to the observations, due to some error in $uWUE$. Most strikingly, the observations show a tendency to move back towards negative values at high VPD, which given the functional form of VPD dependnce, is behavior that is impossible for the theory to capture. This general postive bias the theory explains  the shift towards negative ET reponse observed in the bulk statistics of Section \ref{stats_sec}. The divergence between the theory and the observations for CRO could have been expected. For practical reasons, our theory does not include important sources of variability (particularly between sites) for CRO. We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites and/or years. These deficiencies would explain the inability of our theory to match observations. For example, a superposition of sites with C3 plants (at low enviromental VPD) and C4 plants (at high environmental VPD) would explain observed shift back towards negative ET reponse at high VPD when all sites are binned together, as in Figure \ref{agu_fig}.

The theory for GRA suffers from similiar limitations as for CRO. GRA observations are characterized by a consistent trend back towards negative ET reponse at higher VPD, which the functional form of our theory is incapable of accounting for. As copared to CRO, the divergence between the theory and the observations is even greater for GRA, but these differences are explainable using similiar logic as presented for CRO. GRA will also have differences between sites in C3 vs C4 photosythesis, and irrigation practices, as well as strong seasonal variability in crop height and LAI. As with CRO, the trend back towards negative ET reponse at higher VPD could be explained by a superposition of sites with C3 vs C4 sites. Unfortunately, because detailed site-level data of C3 vs C4 plant type, as well as irrigation practes, was difficult to obtain, we could not bin GRO and GRA further by irrigation and photosyethesis stragtegy. However, we hypothesize that the theory would validate against obsercations much better when photosythesis strategy and irrigation are accounted for.

The comparison between the theory and observations for ENF is unique from the rest of the PFTs. The VPD dependence of the theory, determined by the slope paramter ($g_1$), is not well captured by observations. However, the theory closely captures the x-intercept and VPD$_{crit}$. So, the theory still captures when the ET reponse will be positive or negative, but is less accurate with the magnitude of the reponse. Therefore the theory is still useful for addressing when VPD drives or reduces ET. Also, it's with noting that we used published constants for $g_1$ \citep[from ][]{Lin_2015}, which determines the VPD dependence. We could have fit our own constants so that the theory would best match the observations, but our goal was to use a priori physical constants whenever possible.

While the above discussion shows that our theory has some limitations when applied to some PFTs, especially for crops and grasslands, it does well for DBF and CSH PFTs, and captures the sign of the reponse conistently for ENF. It is useful to also compare our theory, the observed uncertainty in the reponse, and PET, as PET is ofted used as a proxy for ET in hydrometerological studies and drought indicies.  Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5. In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far more negative than 3.5 for the observations, even with all uncertainty included. So the PET reponse to atmospheric dryness is outside the range of all observable responses. This casts doubt on PET's usefulness for capturing the leading order behavior of water budgeting at the land surface. However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ as an alternative to PET to estimating land response and moisture budgeting, and would more accurately account for the response to atmospheric drying..

\subsubsection{Testing VPD$_{crit}$}
As stated in the above discussion of ENF, we are mostly interested in the critical threshold $VPD_{crit}$ for which the sign of the ET reponse to VPD ($\frac{\partial \; ET}{\partial \; VPD}$) is determined. When VPD is less than VPD$_{crit}$, ET reponse to VPD is negative ($\frac{\partial \; ET}{\partial \; VPD}  < 0$, plant response dominates), and when VPD is greater than VPD$_{crit}$, ET reponse to VPD is positive  ($\frac{\partial \; ET}{\partial \; VPD} > 0$ atmospheric demand dominates). Theoretically this $VPD_{crit}$ is only a function of PFT. We can look at how this value of $VPD_{crit}$ matches the data as different observed variables (including uncertainty) are allowed to vary.

Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty ($\sigma$). Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT. Vertical black lines denote where our theory says $VPD_{crit}$ should be. For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates). In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.

Figure \ref{real} generally confirms expectations from Figure \ref{agu_fig}. By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and aerodynamic conductance ($g_a$) on the variability of ET reponse ($\frac{\partial \; ET}{\partial \; VPD}$). Aerodyamic conducntance's scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs. $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}. However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible. The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).

While $g_a$ has a large impact on the magnitude of the scaling, temperature, through Clausius-Clapeyron ($\Delta$), does not appear to have as much of an impact in the complete estimation of ET reponse ($\frac{\partial \; ET}{\partial \; VPD}$). Therefore it appears that the absolute magnitude of the derivative can be mostly attributed to the sign term, and its direct scaling by aerodynamic conductance. Intuitively this is reasonable; aerodynamic conductance will control how dominant balances at the land surface are communicated to the boundary layer. While $\Delta$ controls the efficiency of latent fluxes in carrying heat away from the surface, it appears this is a second order term, relative to $g_a$, for scaling ET reponse.

In general Figures \ref{real} and \ref{agu_fig} match our expectations. Our $\sigma$-based method of uncertainty propagation blurs the idealized expectations the most for GRA and CRO. We therefore have the most confidence in our conclusion based on Equation \ref{d_et} for PFTs CSH, DBF, and ENF, as the sign of the derivative with uncertainty included matches (to leading order) the story when $\sigma$ is held fixed.
